---
title: "Sports Companies Financial Analysis"
author: "Yemi Ajayi"
date: "11-02-2023"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(stringr)
library(openxlsx)
library(lubridate)
library(janitor)
library(glue)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


The main aim of this project is to carry out financial statement analysis on four sportswear brands, covering a time period of 31-12-2011 to 31-12-2020.

The project required me to manipulate the data, carry out data wrangling and create data visualisations in Tableau, to provide information about the brand's business performance and financial position. 

### Load files 

The first thing I done was to load the excel file with all the accounting information

```{r include = FALSE}
Company_Information <- read_excel("~/Sports Company/Company_Information.xlsx")
```

```{r}
print(Company_Information)
```

### Load sheets and format date so they are ready to be used

```{r include=TRUE}
FX <- read_xls(Company_Information, sheet = "FX") %>% 
  mutate(Year = excel_numeric_to_date(Date),
         Year = as.character(Year))
```


You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
